---
title: "Analysis of Patient Data"
Author: Anzhilov Shorena K.
output:
  html_document:
    df_print: paged
Date: r Sys.Date()
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
# Load necessary libraries
library(sqldf)
library(kableExtra)
library(knitr)

# Read in the dataset
df <- read.csv("diabetes_data.csv")
```

```{r data_summary_values, include=FALSE}
# Pre-calculate values for summary paragraph
num_records <- nrow(df)
avg_age <- round(mean(df$Age), 1)
avg_age_female <- round(mean(df$Age[df$Sex == 1]), 1)
avg_age_male <- round(mean(df$Age[df$Sex == 0]), 1)
pct_fruits_veggies <- round(100 * mean(df$Fruits == 1 & df$Veggies == 1), 1)
pct_smokers <- round(100 * mean(df$Smoker == 1), 1)
pct_bmi26_no_activity <- round(100 * mean(df$BMI > 26 & df$PhysActivity == 0), 1)
```

The data set has **`r num_records`** patient records. Males were coded in the data as 0 while females were coded as 1. The mean age for all patients is **`r avg_age`** years (female: **`r avg_age_female`**, male: **`r avg_age_male`**). **`r pct_fruits_veggies`**% of all patients reported eating both fruits and vegetables. **`r pct_smokers`**% of all patients were smokers. **`r pct_bmi26_no_activity`**% of all patients with a BMI over 26 did not report any physical activity.

## Summary of Health Characteristics

```{r avg_age, include=FALSE}
# Average age by gender
avg_age <- sqldf("SELECT Sex, ROUND(AVG(Age), 1) AS avg_age FROM df GROUP BY Sex")
```

```{r pct_diabetes, include=FALSE}
# Percent diagnosed with diabetes by gender
pct_diabetes <- sqldf("
  SELECT Sex, ROUND(100.0 * SUM(Diabetes = 1) / COUNT(*), 1) AS pct_diabetes
  FROM df GROUP BY Sex")
```

```{r pct_chol, include=FALSE}
# Percent with high cholesterol by gender
pct_chol <- sqldf("
  SELECT Sex, ROUND(100.0 * SUM(HighChol = 1) / COUNT(*), 1) AS pct_chol
  FROM df GROUP BY Sex")
```

```{r bmi_stats, include=FALSE}
# Calculate BMI cutoff mean + 2 SD
bmi_mean <- mean(df$BMI, na.rm = TRUE)
bmi_sd <- sd(df$BMI, na.rm = TRUE)
bmi_cutoff <- bmi_mean + 2 * bmi_sd
```

```{r pct_bmi2sd, include=FALSE}
# Percent with BMI > 2 SD above mean by gender
pct_bmi2sd <- sqldf(paste0("
  SELECT Sex, ROUND(100.0 * SUM(BMI > ", bmi_cutoff, ") / COUNT(*), 1) AS pct_bmi2sd
  FROM df GROUP BY Sex"))
```

```{r pct_no_activity, include=FALSE}
# Percent without physical activity by gender
pct_no_phys <- sqldf("
  SELECT Sex, ROUND(100.0 * SUM(PhysActivity = 0) / COUNT(*), 1) AS pct_no_activity
  FROM df GROUP BY Sex")
```

```{r pct_bp, include=FALSE}
# Percent with high blood pressure by gender
pct_bp <- sqldf("
  SELECT Sex, ROUND(100.0 * SUM(HighBP = 1) / COUNT(*), 1) AS pct_bp
  FROM df GROUP BY Sex")
```

```{r merge_results, include=FALSE}
# Merge all statistics by Sex
merged_data <- Reduce(function(x, y) merge(x, y, by = "Sex"),
                      list(avg_age, pct_diabetes, pct_chol, pct_bmi2sd, pct_no_phys, pct_bp))

# Convert Sex to readable gender labels
merged_data$Sex <- ifelse(merged_data$Sex == 0, "Male", "Female")

# Rename columns for display
colnames(merged_data) <- c("Gender", "Avg Age", "% Diabetes", "% High Chol", "% BMI > 2SD", "% No Activity", "% High BP")

# Assign to summary_table for rendering
summary_table <- merged_data
```

```{r show_table, echo=FALSE, results='asis'}
# Display the final table using kableExtra
kable(summary_table, format = "html", caption = "Health Characteristics by Gender") %>%
  kable_styling(position = "left", bootstrap_options = c("striped", "hover"))
```