---
title: "CS5200 Assignment 06.1 - Queriy a Database with SQL in R notebook"
author: "Shorena K. Anzhilov"
date: "Summer 2025, June"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# load libraries
library(DBI)
library(RSQLite)

# Connection to the SQLite database
con <- dbConnect(SQLite(), "OrdersDB.sqlitedb.db")
```

# List tables -- additional for verification.
```{r}
dbListTables(con)
```

--- 1 ---
## Names and phone numbers of all shippers, sorted by name.
```{sql connection=con}
SELECT ShipperName, Phone
FROM Shippers
ORDER BY ShipperName;
```

--- 2 ---
## Number of customers for each country.
```{sql connection=con}
SELECT Country, COUNT(CustomerID) AS NumCust
FROM Customers
GROUP BY Country
ORDER BY Country;
```

--- 3 ---
## The number of products by supplier ID and name, ordered from most to least, restricted to those suppliers who sell at least five products.
```{sql connection=con}
SELECT s.SupplierID, s.SupplierName, COUNT(p.ProductID) AS ProductNum
FROM Suppliers s
JOIN Products p ON s.SupplierID = p.SupplierID
GROUP BY s.SupplierID, s.SupplierName
HAVING COUNT(p.ProductID) >= 5
ORDER BY ProductNum DESC;
```
## Product count for all suppliers this is part of step 3 -- additational. 
```{sql connection=con}
SELECT s.SupplierID, s.SupplierName, COUNT(p.ProductID) AS ProductCount
FROM Suppliers s
JOIN Products p ON s.SupplierID = p.SupplierID
GROUP BY s.SupplierID, s.SupplierName
ORDER BY ProductCount DESC;
```

--- 4 ---
## The most recent order by date, customer's ID, customer name, contact name, and date.
```{sql connection=con}
SELECT c.CustomerID, c.CustomerName, o.OrderDate
FROM Orders o
JOIN Customers c ON o.CustomerID = c.CustomerID
WHERE o.OrderDate = (SELECT MAX(OrderDate) FROM Orders);
```

--- 5 ---
## The total number of unique orders delivered by each shipper, ordered alphabetically by shipper name.

```{sql connection=con}
SELECT s.ShipperName, COUNT(DISTINCT o.OrderID) AS TotalNumOrders
FROM Shippers s
JOIN Orders o ON s.ShipperID = o.ShipperID
GROUP BY s.ShipperName
ORDER BY s.ShipperName;
```
--- 6 ---
## Quinity employees that never worked with a single customer, they appear in the Employees table but not in the Orders table.

```{sql connection=con}
SELECT COUNT(*) AS NumEmployeesWithoutOrders
FROM Employees
WHERE EmployeeID NOT IN (SELECT DISTINCT EmployeeID FROM Orders);
```
--- 7 ----
## Products that contain "Louisiana" at the beginning of the name of the product.
```{sql connection=con}
SELECT COUNT(*) AS LouisianaProductsName
FROM Products
WHERE ProductName LIKE 'Louisiana%';
```

---- 8 ----
## The total number of distinct countries to which at least one order was shipped.

```{sql, connection=con}
SELECT COUNT(DISTINCT Country) AS NumCountriesWithOrders
FROM Customers;
```

---- 9 -----
## Total amount (in terms of revenue) as well as the total number of orders sold by each employee? List the employee name, the total amount sold, and the total number of orders.

```{sql, connection=con}
SELECT e.LastName || ', ' || e.FirstName AS EmployeeName,
       COUNT(DISTINCT o.OrderID) AS TotalOrders,
       SUM(od.Quantity * p.Price) AS TotalRevenueAmount
FROM Employees e
JOIN Orders o ON e.EmployeeID = o.EmployeeID
JOIN OrderDetails od ON o.OrderID = od.OrderID
JOIN Products p ON od.ProductID = p.ProductID
GROUP BY e.EmployeeID
ORDER BY TotalRevenueAmount DESC;
```

--- 10 -----

## Supplier that sells the least number of products but supplies at least one product.
## I am assuming only one suppler here not suppliers so I am using LIMIT 1.

```{sql, connection=con}
SELECT s.SupplierID, s.SupplierName, COUNT(p.ProductID) AS ProductCount
FROM Suppliers s
JOIN Products p ON s.SupplierID = p.SupplierID
GROUP BY s.SupplierID
HAVING COUNT(p.ProductID) >= 1
ORDER BY ProductCount ASC
LIMIT 1;
```

--- 11 ----

## Listing the product name and ID of The product that was ordered most often. 

```{sql, connection=con}
SELECT p.ProductID, p.ProductName, SUM(od.Quantity) AS TotalOrderedNum
FROM OrderDetails od
JOIN Products p ON od.ProductID = p.ProductID
GROUP BY p.ProductID
ORDER BY TotalOrderedNum DESC
LIMIT 1;
```

--- 12 ----

## Listing the product with the product ID and name that that generated the most revenue.

```{sql, connection=con}
SELECT p.ProductID, p.Productname, SUM(od.Quantity * p.Price) AS MostRevenue
FROM OrderDetails od
JOIN Products p ON od.ProductID = p.ProductID
GROUP BY p.ProductID
ORDER BY MostRevenue DESC
LIMIT 1;
```

--- 13 ----

## Listing the total amount spent by all customers who live in either Brazil, Mexico, or Canada. 

```{sql, connection=con}
SELECT SUM(od.Quantity * p.Price) AS TotalAmountSpent
FROM Orders o
JOIN OrderDetails od ON o.OrderID = od.OrderID
JOIN Products p ON od.ProductID = p.ProductID
JOIN Customers c ON o.CustomerID = c.CustomerID
WHERE c.Country IN ('Brazil', 'Mexico', 'Canada');
```

--- 14 ----

## The difference in spending between the country to which the most was sold versus the country to which the least was sold.

```{sql, connection=con}
WITH SpendingByCountry AS (
    SELECT c.Country, SUM(od.Quantity * p.Price) AS Spending
    FROM Orders o
    JOIN OrderDetails od ON o.OrderID = od.OrderID
    JOIN Products p ON od.ProductID = p.ProductID
    JOIN Customers c ON o.CustomerID = c.CustomerID
    GROUP BY c.Country
)
SELECT MAX(Spending) - MIN(Spending) AS SpendingDifference
FROM SpendingByCountry;
```

--- 15 ----

## Which country has the least number of customers.

```{sql, connection=con}
SELECT Country, COUNT(*) AS LeastNumCustomers
FROM Customers
GROUP BY Country
ORDER BY LeastNumCustomers ASC
LIMIT 1;
```

--- 16 ----

## Employee that generated the most revenue (in terms of "dollar amount" sold).

```{sql, connection=con}
SELECT e.EmployeeID, e.LastName || ', ' || e.FirstName AS EmployeeName,
       SUM(od.Quantity * p.Price) AS MostRevenue
FROM Employees e
JOIN Orders o ON e.EmployeeID = o.EmployeeID
JOIN OrderDetails od ON o.OrderID = od.OrderID
JOIN Products p ON od.ProductID = p.ProductID
GROUP BY e.EmployeeID
ORDER BY MostRevenue DESC
LIMIT 1;
```

--- 17 ---- 

## The Listing the Customers (name and ID) that have never bought anything.

```{sql, connection=con}
SELECT CustomerID, CustomerName
FROM Customers
WHERE CustomerID NOT IN (SELECT DISTINCT CustomerID FROM Orders);
```

--- 18 ---

## The average order total per country and List of the country and the average order total.

```{sql, connection=con}
SELECT c.Country AS Country,
       AVG(od.Quantity * p.Price) AS AverageOrderTotal
FROM Customers c
JOIN Orders o ON c.CustomerID = o.CustomerID
JOIN OrderDetails od ON o.OrderID = od.OrderID 
JOIN Products p ON od.ProductID = p.ProductID
GROUP BY c.Country
ORDER BY AverageOrderTotal DESC;
```

--- 19 ---

## The customer that sent the most recent order. Listing of the Order ID, customer name, and country to which the order was shipped.

```{sql, connection=con}
SELECT o.OrderID, c.CustomerName AS CustomerName, c.Country
FROM Orders o
JOIN Customers c ON o.CustomerID = c.CustomerID
WHERE o.OrderDate = (SELECT MAX(OrderDate) FROM Orders);
```

--- 20 ----

## Number of unique suppliers, from how many different (unique) suppliers do the products come from.

```{sql, connection=con}
SELECT COUNT(DISTINCT SupplierID) AS UniqueSuppliersNum
FROM Products;
```

```{r disconnect, include=FALSE}
# Step: Disconnect from the database
dbDisconnect(con)














